%\documentstyle[12pt]{report}
%\usepackage{lscape}
%\setlength{\parindent}{0mm}
%\setlength{\parskip}{6pt}
%\renewcommand{\baselinestretch}{1.5}
%\setlength{\topmargin}{0pt}
%\setlength{\headheight}{0pt}
%\setlength{\headsep}{0pt}
%\setlength{\footheight}{0pt}
%\setlength{\footskip}{45pt}
%\setlength{\textwidth}{430pt}
%\setlength{\textheight}{660pt}
%\setlength{\oddsidemargin}{10pt}
%\begin{document}

\section*{\begin{center} ACKNOWLEDGEMENTS \end{center}}
It  would be impossible to list all those who have provided me with vital
help. But I must mention those to whom I owe a special debt.

I am particularly grateful to my supervisor Prof F M Mahomed for his
unfailing  support throughout the writing of this thesis. His multiple
advices and suggestions  have considerably enhanced the readability of
earlier drafts. Foremost, his enthusiasm and dedication to research have
been my source of inspiration.

I have also benefited from the most needed moral support of my family and
friends. Their phone calls and letters have kept me from stress and its
consequences.

I would like to register my sincere appreciation to Prof N H Ibragimov and
Dr C Qu. The former  pointed out the reference Lie--Engel (1893, chapter 7)
to me and the latter
collaborated to Chapter 2 during his visit to the Department of Computational
and Applied Mathematics of the University of the Witwatersrand.

Finally, I am very much indebted to the University of the Witwatersrand for
its financial assistance through the Postgraduate Merit Award and the
ASA Educational Trust Postgraduate Award. I also received
a substantial amount of money courtesy of a research grant. For this last
grant, I would  like to thank Prof F M Mahomed who made it possible. 

%\end{document}
