%\documentclass[12pt]{report}
%\usepackage{lscape}
%\setlength{\parindent}{0mm}
%\setlength{\parskip}{6pt}
%\renewcommand{\baselinestretch}{1.5}
%\setlength{\topmargin}{0pt}
%\setlength{\headheight}{0pt}
%\setlength{\headsep}{0pt}
%\setlength{\footheight}{0pt}
%\setlength{\footskip}{45pt}
%\setlength{\textwidth}{430pt}
%\setlength{\textheight}{660pt}
%\setlength{\oddsidemargin}{10pt}

%\begin{document}

\chapter*{\begin{center} References \end{center}}
\addcontentsline{toc}{chapter}{\protect\numberline{}{References}}
 
Abel N H 1839 {\em Collected Papers} (in French), by
B. Holmboe (Olso).
 
Anderson R L, Hanard J and Winternitz P 1982
Systems of ordinary differential equations with nonlinear 
superposition principles, {\em Physica {\bf4D}}, 164--182.

Anderson R L and Ibragimov N H 1979 {\em Lie--B\"acklund Transformations in
Applications} (Philadelphia: SIAM).

Belinfante J G, Kolman B and  Smith A H 1966, 1968
An introduction to Lie groups and Lie algebras with applications,
{\em SIAM Review} Vol. 8, No. 1, 11--45 and Vol 10, No. 2, 160--195.

Bluman G W and Kumei S 1989 {\em
Symmetries and Differential Equations} (New York: Springer--Verlag).

Campbell J E 1903 {\em Theory of Continuous Groups} (Oxford:
Oxford Univ. Press).

Chern S 1937 Sur la g\'eometrie d'une \'equation differentielle du 3e ordre,
{\em C. R. Acad. Sci. Paris}, 1227.

Conrad M M,  Mahomed F M and Leach P G L 1994  Solution of generalised
Emden-Fowler equations with two symmetries, {\em Int. J. Non--Linear
Mechanics} {\bf 29}, 529.

Damianou P A and Sophocleous C 1999 Symmetries of Hamiltonian systems with
two degrees of freedom, {\em J. Math. Phys.} {\bf 40}(1), 210--235.

Dickson L E 1924 Differential Equations from a group standpoint, {\em Ann. Math.} {\bf 25}, 287.

Engel F 1899 Sophus Lie, {\em Berichte \"uber die Verhandlungen der
K\"oniglich S\"achsischen Gesellschaft de Wissen zu Leipzig,
Math.--physische Klasse} {\bf 51}.

Einsenhart L P 1933 {\em Continuous Groups of
Transformations} (Princeton: Princeton Univ. Press).

Ermakov V 1880 {\em Univ. Isz. Kiev Series} III {\bf 9}, 1
(trans. by O Harin).

Forsyth A 1921 {\em Treatise on Differential Equations}, 5th ed.
(London: Macmillan and Co.).

Fritzsche B 1999 Sophus Lie: A sketch of his Life and work,
{\em  J. of Lie Theory} {\bf 9}, 1--38.  

Gat O 1992  Symmetry algebras of third--order ordinary
differential equations, {\em J. Math. Phys.} {\bf 33}, 2966--2971.

Gonz\'alez--Gasc\'on F and Gonz\'alez--L\'opez  A 1983
Symmetries of differential equations IV, {\em J. Math. Phys.} {\bf 24},
2006--2021.

Gorringe V M and Leach P G L 1988
Lie Point symmetries of second--order linear ordinary differential
equations, {\em Quaetiones Math.} {\bf 11}, 95--117.

Herlt H and  Stephani H 1992  Invariant transformations of the class
$y''=F(x)y^2$ of differential equations, {\em J. Math. Phys.} {\bf 33}, 3983. 

Herlt E 1996 Spherically symmetric nonstatic perfect fluid solutions with
shear, {\em Gen. Relativity and Grav.} {\bf28}(8), 919.

Hietarinta J 1987 Direct methods for the search of the second invariant,
{\em Phys. Rep.} {\bf 147}, 87.

Hill J M  1992 {\em Differential Equations and Group Methods for Scientists
and Engineers} (Boca Raton, Florida: CRC Press).

Ibragimov N H 1985 {\em Transformation Groups Applied to Mathematical
Physics} (Dordrecht: D Reidel).

Ibragimov N H  1994, 1995, 1996 {\em CRC Handbook of Lie Group
Analysis of Differential Equations} (Boca Raton, Florida: CRC Press).

Ibragimov N H 1995  English trans. of Lie 1881, in
CRC Handbook of Lie group analysis of differential equations, vol. 2
(Boca Raton, Florida: CRC Press).

Ibragimov N H and Mahomed F M 1996  Ordinary differential equations,
in  CRC Handbook of Lie group analysis and differential equations,
chap. 8, vol. 3 ( Boca Raton: CRC Press).

Ibragimov N H, Kara A H and Mahomed F M 1998 Lie-B\"acklund and
Noether symmetries with applications, {\em Nonlinear Dynamics} {\bf 15}, 115.

Ibragimov N H 1999 {\em Elementary Lie Group Analysis
and Ordinary Diffenrential Equations} (New York: Wiley and Sons).

Ibragimov N H, Khalique C M and Mahomed F M 1999
All linear ordinary differential equations admitting  contact symmetries,
Proceedings of the International Conference at the Sophus Lie Centre, Mars
Pub., Symmetri Foundation.

Ince E L 1927 {\em Ordinary Differential Equations} (London: Longmans).

Jacobson N 1955 {\em Lie Algebras} (New York: Interscience).

Kamke E 1983 {\em Differentialgleichungen -- L\"osungsmethoden
und L\"osungen I.} (Stuttgart: B G Teubner).

Kara A H, Vawda F E and Mahomed F M  1994  Symmetries of first integrals and
solutions to differential equations, {\em Lie Group and their Applications}
Vol. {\bf 1} No. 2, 27--48

Kowalewski G 1950 {\em Bestand und Wandel} (M\"unchen: R Oldenbourg).

Kramer D, Stephani H, MacCallum M and Herlt E 1980 {\em Exact Solutions of
Einstein's Field Equations} (Cambridge: Cambridge University Press).

Krause J and Michel L 1988 \'Equations Diff\'erentielles
Lin\'eaires d'ordre $n>2$ ayant une alg\`ebre de Lie de symmetrie
de dimension $n+4$, {\em C. R. Acad. Sci.} {\bf 307} S\'erie I,
905--910.

Kustaanheimo P and Qvist B 1948  A note on some general solutions of the
Einstein field equations in a spherically symmetric world, {\em Soc. Sci.
Fennica, Comm. Phys.--Math.} {\bf XIII} 12.

Leach  P G L, Maartens R and Maharaj S D 1992 Self--similar solutions of the
generalized Emden--Fowler equation, {\em Int. J. Non--Linear Mech.}, Vol.
{\bf27} No. 4, 575--582. 

Levi--Civita T and Amaldi U 1927 {\em Lezioni di
meccanica razionale}, vol. 2, part 2, 452--469 (Bologna: Zanichelli Bologna). 

Lie S 1871 {\em Over en Classe geometriske
Transformationer}, doctoral thesis, University of Christiana, Norway.

Lie S 1874 Begr\"undung einer Invariantentheorie der
Ber\"uhrungstransformationen, Math. Ann. {\bf 8}, 215--288.

Lie S 1881  \"Uber die Integration durch bestimmte
Integrale von einer Klasse linear partialler Differentialgleichungen,
{\em Arch. for Math.}, {\bf 6}, 328--368.

Lie S 1883  Klassification und Integration von
gew\"ohneichen Differentialgleichungen zwischen $x,\;y$, die eine Gruppe
von Transfomationen gestatten, {\em Arch. for Math.} VIII ({\bf 4}), 328.

Lie S 1884  Differential invariants, trans. by
M Ackerman 1976, comments by R Hermann 1976, in Lie group: History,
Frontiers and Applications, vol. 3, (Brookline: Math. Sc. Press). 

Lie S and Engel F 1888, 1890, 1893 {\em Theorie der
Transformationsgruppen} (Leipzig: B G Teubner).

Lie S 1889 Die infinitesimalen Ber\"uhrungstransformationen
der Mechanik, S\"achs. Akad. d. wiss, Berlin, vol. 41, 145--156.

Lie S 1891 {\em Differentialgleichungen} (New York: Chelsea, reprinted 1967).

Lie S and  Scheffers G 1891 {\em Vorlesugen \"uber
Differentialgleichungen mit bekanten infinitesimalen Transformationen}
(Leipzig: B G Teubner, reprinted 1912).


Lie S and Scheffers G 1896
{\em  Geometrie der Ber\"uhrungstransformationen} (Leipzig: B G Teubner).

Lie S  1922 {\em Gesammelte Abhandlungen}, vol. 3 (Leipzig: B G Teubner).

Logan J D  1977 {\em Invariant Variational Principles}
(New York: Academic Press).

Lutzky M 1979 Dynamical symmetries and conserved quantities,
{J. Phys. A}{\bf 11}, 249--258.

Maharaj S D, Leach P G L and Maartens R 1996 Expanding
spherically symmetric models without shear, {\em Gen. Relativity and Grav.}
{\bf28}(1), 35.

Mahomed F M and  Leach P G L 1988 Normal forms for third--order equations,
{\em Proceedings of the Workshop on Finite--dimensional Integrable
Nonlinear Dynamical Systems}, P G L Leach and W H Steeb edd
(Singapore: World Science Publications).

Mahomed F M 1989 {\em Symmetry Lie Algebras of $n$th
Order Ordinary Differential Equations}, PhD thesis, Univ. of the
Witwatersrand, Johannesburg.

Mahomed F M and Leach P G L 1989  Lie algebras associated with scalar
second--order ordinary differential equations, {\em J. Math. Phys.}
{\bf 30}(12), 2770--2777.

Mahomed F M and Leach P G L 1990
Symmetry Lie algebras of $n$th order ordinary differential equations,
{\em J. Math. Anal. Appl.} {\bf 151}, 80--106.

Mahomed F M and Leach P G L 1991
Contact symmetry algebras of scalar
second--order ordinary differential equations, {\em J. Math. Phys.} {\bf 32},
2951--2055.

Mubarakzyanov GM  1963 {\em Izv. Vysshikh Uchebn. Zavedeni\u i Mat.}
No. 1 ({\bf 32}), 114; No. 3 ({\bf 34}), 99; No. 4 ({\bf 35}),
104.

Noether E 1918 Invariante Variationprobleme, {\em Nachrichten der
Wissenschaften in G\"ottingen Mathematish--Physikalische Klasse} {\bf 2},
235--257. (English translation in {\em Transport Theory and Statistical
Physics} {\bf 1}(3), 1971, 186--207).

Olver P J 1986 {\em Applications of Lie Groups to Differential Equations}
(New York: Springer--Verlag). 

Olver P J 1994  Differential invariants and invariant
differential equations,{\em  Lie Group and their Applications}, vol. 1,
July--August, 177--192.

Ovsiannikov L V 1962 {\em Group Properties of Differential Equations}
(Novorsibirsk: USSR Acad. of Sci., Siberian Branch).

Ovsiannikov L V 1982 {\em Group Analysis of Differential
Equations} (New York: Academic Press).

Patera J and Winternitz P  1977
Subalgebras of real three and four dimensional Lie algebras,{\em  J. Math.
Phys.} {\bf 18}, 1449--1455.
 
Petrov A Z 1966 {\em New methods in General Relativity}
(in Russian) (Moscow: Nauka).

Polyanin A D and Zaitsev V F  1995 {\em Handbook of Exact Solutions for Ordinary
Differential Equations} (Boca Raton: CRC Press).

Purkert W 1984  Zum Verh\"altnis von Sophus Lie und Friedrich Engel,
{\em Wissenschaftl. Zeitschr. d. Ernst--Moritz--Arndt--Universit\"at
Greifswald, Math.--naturwissenschaftl. Reihe} {\bf 33}, 30.

Ruban V A 1969 {\em ZhETF} {\bf5b}, 1914 [{\em Sov. Phys. JETP} {\bf29},1027].

Sarlet W and Cantrijn F 1981  Generalizations of Noether's theorem in
classical mechanics, {\em SIAM Rev.} {\bf 23}, 467--494.

Schwarz F 1983 Contact symmetries of the harmonic
oscillator, {\em J. Phys. A} {\bf 16} L133.

Sen T 1987  PhD thesis, State University of New York, Stony Brook.

Stephani H 1989 {\em Differential Equations--Their Solutions
Using Symmetries} (Cambridge: Cambridge University Press). 

Stephani H and Wolf T 1996 Spherically
symmetric perfect fluids in shear--free motion---the symmetry approach,
{\em Class. Quantum Grav.} {\bf 13}, 1261--1271. 

Svirshchevskii S R 1993 Lie--B\"acklund symmetries
of linear ODEs and invariant linear spaces, in Modern Group Analysis, ed.
G N Yakovenko, Moscow.

Svirshchevskii S R 1995 Lie--B\"acklund symmetries
of linear ODEs and generalized separation of variables in nonlinear equations,
{\em Phys. Lett. A} {\bf 199}, 344--348.

Tresse A R 1894 Sur les invariants diff\'erentiels
des groupes continus de transformations, {\em Acta Math.} {\bf 18}, 1--88.

Tresse A R 1896 {\em D\'etermination des invariants
ponctuels de l'\'equation du second ordre $y''=\omega (x,y,y')$}
(Leipzig: S Hirzel).

Vessiot E 1906  Sur l'int\'egration m\'ecanique
des transformations de contact infinit\'esimals, {\em Soc. Math. de France},
Bull., vol. 34, 230--269.

Wafo Soh C, Mahomed F M and Qu C 1999a Contact symmetries of scalar ordinary
differential equations, submitted to {\em Quaetiones Math}.

Wafo Soh C and Mahomed F M 1999b Noether symmetries of $y''=f(x)y^n$
with applications to non--static spherically symmetric perfect fluid
solutions, {\em Class. and Quantum Grav.} {\bf 16}, 3553--3566.

Wafo Soh C and Mahomed F M 1999c Preliminary group classification of
$y''=f(x)y^2$, {\em Math. and Comp. Appl.} {\bf 3} (4), 289--295.

Wafo Soh C and Mahomed F M 1999d Generalization of the Abel--Forsyth
Formula: a symmetry approach, submitted to {\em Nonlinear Dynamics}.

Wafo Soh C and Mahomed F M 1999e Symmetry breaking for a system of two
linear ordinary differential equations, to appear in
{\em Nonlinear Dynamics}.

Wafo Soh C and Mahomed F M 1999f Linearization of systems of nonlinear
differential equations, submitted to {\em Int. J. of Nonlinear Mechanics}.

Wafo Soh C and Mahomed F M 1999g Canonical forms of systems of two
second--order ordinary differential equations admitting symmetry Lie
algebras, submitted to {\em  J. Math. Phys.}.

Whittaker E T 1927 {\em Analytical dynamics} (Cambridge: Cambridge
Univ. Press).

Wyman M 1976  Jeffrey--Williams Lecture: Nonstatic radially symmetric
distributions of matter {\em Can. Math. Bull.} {\bf 19}, 343. 

Yumaguzhin V A 1997  Contact classification of
3--Order linear ODEs, {\em The Diffeity Institute Preprint Series},
http://ecfor.rssi.ru/\~{ } diffeity.

%\end{document}
